Biotic invasions fuck with the manageability of pollination networks

### 2017-02-24 paper/draft_2.Rmd
Cosmetic changes
Change dependency to dependence troughout

**E. Fernando Cagua^1\*^, Kate L Wootton^1,2§^, Daniel B. Stouffer^1#^**

@@

To bridge this gap, we outline an approach to apply theory on the control of complex systems in an ecological context and implement them using empirical data.
Specifically, we use a set of ten pairs of uninvaded and invaded plant-pollinator communities to investigate the link between invasion, network structure and ecological management. 
While doing so, we focus on two particular questions.
First, grounded on the difficulties usually involved with invasive species eradication and ecosystem restoration [@Woodford2016], we ask whether invaded networks have lower levels of "manageability" than their uninvaded counterparts; that is, whether they require a greater proportion of species to be managed to achieve the same level of control. 
Second, we ask whether some species are more important than others at driving the population dynamics of the community and which factors determine this importance. 
We focus on these two particular applications for a variety of reasons. 
First, biotic invasions are known to produce tractable changes in the structure of ecological networks, and these changes can be particularly pronounced in mutualistic networks between plants and pollinators where biotic invasions have been shown to modify the strength of species interactions and the degree of network nestedness and connectivity [@Olesen2002; @Aizen2008; @Bartomeus2008; @Vila2009; @Traveset2013]. 
Second, plant-pollinator networks provide an ideal framework to answer these questions. 
On the one hand, community networks that quantify relative levels of interaction are readily available. 
On the other, the bipartite nature of pollination networks makes it possible to simplify assumptions of how these interactions translate into interspecific effects.

@@

We then examined whether any species-level structural properties can predict our metrics of species importance—the frequency to which a species was a driver or a superior node (*f~D~* and *f~S~*, respectively).
We used a set of generalised linear mixed-effects models (with binomial error structure) with the relative frequencies as the response variables. 
As predictors in this model, we included measures of centrality (degree and eigen-centrality), which have been found to be strong predictors of importance in a coextinction context [@Memmott2004]; a measure related to network robustness (contribution to nestedness), as nestedness has been proposed as one of the key properties that promote stability in mutualistic networks [@Saavedra2011]; and two measures of strength of association and dependence (visitation and dependence strength), as their distribution determines the extent of interspecific effects [@Bascompte2006].
To facilitate comparison among the continuous variables we scaled so that they all had a mean of zero and a standard deviation of one.
In addition, we also included guild (plant or pollinator), and whether the species is invasive or not, as categorical fixed effects.
Although the importance of plants and pollinators or invasive and non-invasive species could respond differently to our structural metrics, our data set did not contain enough variation to include the corresponding interactions terms for these latter two predictors. 
All network metrics were calculated using the R package bipartite 2.06 [@Dormann2008]. 
Lastly, we allowed for variation between different communities by including the network identity as a random effect [@Bates2015]. 
Candidate models and estimates were assessed using the same procedure as in the *n~D~* models.

@@

Invasive species were classified as superior nodes *and* driver nodes in every single network they were present; that is, they always had the highest relative *f~S~* and *f~D~* (Figure 5a). 
The model results suggest that this differences among invasive and native species are underpinned, not by any intrinsic property of the invasive species but instead, by species properties that apply to invasive and native species alike (Table 1). 
Specifically, we found that a species is more likely to be classified as a superior node if it had a large dependence strength (its sum of the dependences). 
To a smaller extent, visitation strength (its sum of visits) and the species degree also had a positive relationship with $f_S$. 
In contrast, the relationships between species structural properties and $f_D$ were more convoluted (Table 1). 
Both invasive species (which for example tend to have a large degree, and high dependence strength) and pollinators (which generally have a smaller degree and dependence strength) were classified as driver nodes in a large proportion of the considered matchings.

@@

```{r fig-dependency-fs, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 3.245, fig.height = 3.2, fig.cap = "**Relatiosnships between _f~S~_ and _f~D~_ and species' structural properties**. (a) In all networks were they were present, invasive species were classified as superior (_f~S~_) and driver (_f~D~_) nodes in all possible control configurations. (b) Dependency strength is the single most important factor explaining _f~S~_. Visitation strength and  degree also had an important albeit comparatively smaller effect (dashed lines correspond to ± one standard deviation on these factors). Invasive species are depicted as circles."}
data <- mo 
trans <- function(x, orig){
  x * attr(orig, "scaled:scale") + attr(orig, "scaled:center")
}

@@

p1 <- pred_data %>%
  ggplot() +
  geom_line(aes(x = de, y = p, linetype = type, colour = type, group = interaction(n, vi)), show.legend = F, size = 0.5) +
  scale_x_log10(expand = c(0,0)) +
   geom_point(data = dplyr::arrange(data_pp, invasive) %>% dplyr::filter(invasive == T), aes(x = de_s, y = pred_ri, fill = invasive, shape = guild, size = invasive), show.legend = F) +
    scale_shape_manual(values = c(21, 23)) +
  scale_size_manual(values = c(1.5,1.5)) +
  scale_fill_manual(values = c("white","#e41a1c")) + 
  scale_color_manual(values = c("#525252", "black")) +
  scale_linetype_manual(values = c(2, 1)) +
  xlab("dependency strength") +
  ylab(expression(f[S])) +
  fer_theme

@@

```{r table-species-model-coef, echo = FALSE, warning = FALSE, message = FALSE, results = "asis", sanitize = FALSE, cache = T}
coef_tables <- mod %>%
	plyr::llply(function(x){
		x %>%
			coef.glmulti(select = 0.95) %>%
  as.data.frame()  %>%
  dplyr::add_rownames("factor") %>%
  dplyr::mutate(pare = grepl("(", factor, fixed =T),
                factor = plyr::mapvalues(factor, 
                                         c("de", "n", "vi", "eig_cen", 
                                           "guildpol", "d", "invasiveTRUE"),
                                         c("dependency strength", 
                                           "contribution to nestedness",
                                           "visitation strength",
                                           "eigen-centrality", 
                                           "guild (pollinator)", "degree", "invasive sp."))) %>% 
  magrittr::set_names(c("factor", "est.", "var", "mod.", "imp.", "C.I.", "pare")) %>%
			dplyr::mutate(absEstimate = abs(est.)/C.I.) %>%
  dplyr::arrange(-pare, -imp., -absEstimate, factor) %>% 
  dplyr::select(-pare, - absEstimate)
	})

@@

Contrary to our initial hypothesis, we found some evidence that invaded communities might be easier to manage than uninvaded ones. 
Our results reveal, however, that this effect is comparatively small and the structural differences among different networks are more importantly related to potential differences in our ability to alter the state of the community.
Despite the small effect of invasion status at a network level, we found that invasive mutualists occupy a particularly dominant role in their communities. 
Not only changes on their abundance have the potential to propagate broadly trough the community and in turn, affect the abundances of many other species, but they are also indispensable to fully control the plant-pollinator network. 
At a community level, we demonstrate that the manageability of mutualistic networks is strongly governed by the asymmetric nature of mutual dependencies—which constitute the foundations of the structure and stability of mutualistic networks [@Memmott2004; @Vazquez2004; @Bascompte2006; @Lever2014; @Astegiano2015]. 
Moreover, these mutual dependences seem to be importantly constrained by the concerted effects of both, the patterns of species richness at each trophic guild, and the species' degree distribution [@Melian2002; @Bluthgen2007]. 
Indeed, the difference between the driver node density (*n~D~*) of our empirical networks and that of randomisations depended strongly on the evaluated null model. 
While the empirical *n~D~* was indistinguishable from that of networks with a random structure that maintained the degree or strength of each species in the community, it was much larger than that of randomisations in which the directed network was unchanged but where the observed patterns of dependence were broken.

@@

Despite its conceptual advantages, our approach still relies on knowledge of the network of interaction between species. 
Nevertheless, our results and previous basic research show that as long as the proportion of sampled links is enough to provide indication of the actual degree distribution (and in turn the species dependences), *n~D~*, *f~S~* and *f~D~* can still provide a simple, straightforward, and theoretically-informed indication of the degree to which the community is self-regulated and therefore how difficult it might be to modify its state in one way or another using some of the species that integrate it [Supporting Information S5; @Nepusz2012].

@@

*n~D~*
  ~ The density of driver nodes

@@

Maximum matching
  ~ A matching with the largest possible matching size *and* matching weight.

Superior node
  ~ The node at the start of a matched link. From the control perspective, superior nodes make up the chain(s) that propagate the control signal(s) through the network.

>>>>

**E. Fernando Cagua^1\*^, Kate L. Wootton^1,2§^, Daniel B. Stouffer^1#^**

@@

To bridge this gap, we outline an approach to apply theory on the control of complex systems in an ecological context and implement them using empirical data.
Specifically, we use a set of ten pairs of uninvaded and invaded plant-pollinator communities to investigate the link between invasion, network structure and ecological management. 
While doing so, we focus on two particular questions.
First, grounded on the difficulties usually involved with invasive species eradication and ecosystem restoration [@Woodford2016], we ask whether invaded networks have lower levels of "manageability" than their uninvaded counterparts; that is, whether they require a greater proportion of species to be managed to achieve the same level of control. 
Second, we ask whether some species are more important than others at driving the population dynamics of the community and which factors determine this importance. 
We focus on these two particular applications for a variety of reasons. 
First, biotic invasions are known to produce tractable changes in the structure of ecological networks, and these changes can be particularly pronounced in mutualistic networks between plants and pollinators where biotic invasions have been shown to modify the strength of species interactions and the degree of network nestedness and connectivity [@Olesen2002; @Aizen2008; @Bartomeus2008; @Vila2009; @Traveset2013]. 
Second, although species are involved in multiple types of interactions, plant-pollinator networks provide an ideal framework to answer these questions. 
On the one hand, community networks that quantify relative levels of interaction are readily available. 
On the other, the bipartite nature of pollination networks makes it possible to simplify assumptions of how these interactions translate into interspecific effects.

@@

We then examined whether any species-level structural properties can predict our metrics of species importance—the frequency to which a species was a driver or a superior node (*f~D~* and *f~S~*, respectively).
We used a set of generalised linear mixed-effects models (with binomial error structure) with the relative frequencies as the response variables. 
As predictors in this model, we included measures of centrality (degree and eigen-centrality), which have been found to be strong predictors of importance in a coextinction context [@Memmott2004]; a measure related to network robustness (contribution to nestedness), as nestedness has been proposed as one of the key properties that promote stability in mutualistic networks [@Saavedra2011]; and two measures of strength of association and dependence (visitation and dependence strength, this is the sum of visits a species receives or performs, and the sum of dependences of all species on the focal species, respectively), as their distribution determines the extent of interspecific effects [@Bascompte2006].
To facilitate comparison among the continuous variables we scaled so that they all had a mean of zero and a standard deviation of one.
In addition, we also included guild (plant or pollinator), and whether the species is invasive or not, as categorical fixed effects.
Although the importance of plants and pollinators or invasive and non-invasive species could respond differently to our structural metrics, our data set did not contain enough variation to include the corresponding interactions terms for these latter two predictors. 
All network metrics were calculated using the R package bipartite 2.06 [@Dormann2008]. 
Lastly, we allowed for variation between different communities by including the network identity as a random effect [@Bates2015]. 
Candidate models and estimates were assessed using the same procedure as in the *n~D~* models.

@@

Invasive species were classified as superior nodes *and* driver nodes in every single network they were present; that is, they always had the highest relative *f~S~* and *f~D~* (Figure 5a). 
The model results suggest that this differences among invasive and native species are underpinned, not by any intrinsic property of the invasive species but instead, by species properties that apply to invasive and native species alike (Table 1). 
Specifically, we found that a species is more likely to be classified as a superior node if it had a large species strength (the sum of the dependences of other species on the species of interest). 
To a smaller extent, visitation strength (its sum of visits) and the species degree also had a positive relationship with $f_S$. 
In contrast, the relationships between species structural properties and $f_D$ were more convoluted (Table 1). 
Both invasive species (which for example tend to have a large degree, and high dependence strength) and pollinators (which generally have a smaller degree and dependence strength) were classified as driver nodes in a large proportion of the considered matchings.

@@

```{r fig-dependency-fs, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 3.245, fig.height = 3.2, fig.cap = "**Relatiosnships between _f~S~_ and _f~D~_ and species' structural properties**. (a) In all networks were they were present, invasive species were classified as superior (_f~S~_) and driver (_f~D~_) nodes in all possible control configurations. (b) Dependence strength is the single most important factor explaining _f~S~_. Visitation strength and  degree also had an important albeit comparatively smaller effect (dashed lines correspond to ± one standard deviation on these factors). Invasive species are depicted as circles."}
data <- mo 
trans <- function(x, orig){
  x * attr(orig, "scaled:scale") + attr(orig, "scaled:center")
}

@@

p1 <- pred_data %>%
  ggplot() +
  geom_line(aes(x = de, y = p, linetype = type, colour = type, group = interaction(n, vi)), show.legend = F, size = 0.5) +
  scale_x_log10(expand = c(0,0)) +
   geom_point(data = dplyr::arrange(data_pp, invasive) %>% dplyr::filter(invasive == T), aes(x = de_s, y = pred_ri, fill = invasive, shape = guild, size = invasive), show.legend = F) +
    scale_shape_manual(values = c(21, 23)) +
  scale_size_manual(values = c(1.5,1.5)) +
  scale_fill_manual(values = c("white","#e41a1c")) + 
  scale_color_manual(values = c("#525252", "black")) +
  scale_linetype_manual(values = c(2, 1)) +
  xlab("dependence strength") +
  ylab(expression(f[S])) +
  fer_theme

@@

```{r table-species-model-coef, echo = FALSE, warning = FALSE, message = FALSE, results = "asis", sanitize = FALSE, cache = T}
coef_tables <- mod %>%
	plyr::llply(function(x){
		x %>%
			coef.glmulti(select = 0.95) %>%
  as.data.frame()  %>%
  dplyr::add_rownames("factor") %>%
  dplyr::mutate(pare = grepl("(", factor, fixed =T),
                factor = plyr::mapvalues(factor, 
                                         c("de", "n", "vi", "eig_cen", 
                                           "guildpol", "d", "invasiveTRUE"),
                                         c("species strength", 
                                           "contribution to nestedness",
                                           "visitation strength",
                                           "eigen-centrality", 
                                           "guild (pollinator)", "degree", "invasive sp."))) %>% 
  magrittr::set_names(c("factor", "est.", "var", "mod.", "imp.", "C.I.", "pare")) %>%
			dplyr::mutate(absEstimate = abs(est.)/C.I.) %>%
  dplyr::arrange(-pare, -imp., -absEstimate, factor) %>% 
  dplyr::select(-pare, - absEstimate)
	})

@@

Contrary to our initial hypothesis, we found some evidence that invaded communities might be easier to manage than uninvaded ones. 
Our results reveal, however, that this effect is comparatively small and the structural differences among different networks are more importantly related to potential differences in our ability to alter the state of the community.
Despite the small effect of invasion status at a network level, we found that invasive mutualists occupy a particularly dominant role in their communities. 
Not only changes on their abundance have the potential to propagate broadly trough the community and in turn, affect the abundances of many other species, but they are also indispensable to fully control the plant-pollinator network. 
At a community level, we demonstrate that the manageability of mutualistic networks is strongly governed by the asymmetric nature of mutual dependences—which constitute the foundations of the structure and stability of mutualistic networks [@Memmott2004; @Vazquez2004; @Bascompte2006; @Lever2014; @Astegiano2015]. 
Moreover, these mutual dependences seem to be importantly constrained by the concerted effects of both, the patterns of species richness at each trophic guild, and the species' degree distribution [@Melian2002; @Bluthgen2007]. 
Indeed, the difference between the driver node density (*n~D~*) of our empirical networks and that of randomisations depended strongly on the evaluated null model. 
While the empirical *n~D~* was indistinguishable from that of networks with a random structure that maintained the degree or strength of each species in the community, it was much larger than that of randomisations in which the directed network was unchanged but where the observed patterns of dependence were broken.

@@

Despite its conceptual advantages, our approach still relies on knowledge of the network of interaction between species. 
We do not argue that management decisions should be based on a single technique. 
Nevertheless, our results and previous basic research show that as long as the proportion of sampled links is enough to provide indication of the actual degree distribution (and in turn the species dependences), *n~D~*, *f~S~* and *f~D~* can still provide a simple, straightforward, and theoretically-informed indication of the degree to which the community is self-regulated and therefore how difficult it might be to modify its state in one way or another using some of the species that integrate it [Supporting Information S5; @Nepusz2012].

@@

@@

Maximum matching
  ~ A matching with the largest possible matching size *and* matching weight. 
Superior node
  ~ The node at the start of a matched link. From the control perspective, superior nodes make up the chain(s) that propagate the control signal(s) through the network.

### 2017-02-24 paper/supp_info_2.Rmd
Change dependency to dependence troughout

Visitation frequency has been shown to be an appropriate surrogate for inter-specific effects in pollination networks [@Vazquez2005; @Bascompte2006]. 
Nevertheless visitation is not equivalent to pollen deposition and might be insufficient to reflect the dependencies of plants on animals and vice versa [@Alarcon2010; @King2013]. 
We therefore investigated the effect of calculating the dependencies using visitation or pollination effectiveness and importance---two metrics more proximate to plant reproductive success (Figure \@ref(fig:fig-visitation-vs-deposition-vs-importance)).
We did this by comparing *(i)* the manageability of the community and *(ii)* the percentage of interactions that maintained the direction of dependency. 
To do that, we used data collected by @Ballantyne2015 from a low diversity pollination community at a dry lowland heathland in Dorset, UK  (50° 43.7'N 2° 07.2'W). 
First, deposition networks were quantified using the mean single visit deposition---the number of conspecific pollen grains effectively deposited on a virgin stigma during a single visit by a particular animal [@NeEman2010; @King2013; @Ballantyne2015]. 
Second, visitation networks were constructed counting the visits to flowers during Single Visit Depositions. 
Finally, pollinator importance networks were constructed as the product of pollinator efficiency and visit frequency.

@@

The choice of weighting used can also have an impact on at the species level.
Therefore we calculated *f~D~* and *f~S~* (the frequency at which each species is classified as a driver or superior node, respectively, within the set of all possible matchings in a network) and calculated it's correlation among all three weighting schemes.
Although visitation and deposition produce moderately different results, we found a very strong agreement between the order produced by visitation and importance (Table S2). 
Finally, we also investigated whether the asymmetry of mutual dependency, which defines the direction of control, was consistent among the three possible weighting schemes. We found again that the direction of the dominant dependency was maintained was consistent for 95\% of the interactions weighted by visitation or importance, the two most appropriate metrics for pollinator and plant dependency (Table S2).

@@

mapply(my_paste, correlations$r, correlations$P) %>%
  matrix(ncol = length(dir_net_b)) %>% magrittr::set_colnames(nam) %>%
  magrittr::set_rownames(nam) %>% `[<-`(lower.tri(.), lt) %>% 
  xtable::xtable(caption = "\\textbf{Agreement among network weighting schemes}. Spearman correlation coeffcients (with p-value) of species' $f_D$ and the percentage of interactions that share the direction of dependency obtained using the three weighting schemes and an unweighted scheme.") %>%
	print(comment = FALSE, include.rownames = TRUE, caption.placement = "top", booktabs = T)

@@

Altogether, evidence supports the idea that visitation is a suitable metric to estimate the mutual dependency of species pairs. First, it is directly related to pollinator foraging. Second, it produces results consistent, at least within our controllability framework, with plant reproductive success (as estimated by the importance metric).

>>>>

Visitation frequency has been shown to be an appropriate surrogate for inter-specific effects in pollination networks [@Vazquez2005; @Bascompte2006]. 
Nevertheless visitation is not equivalent to pollen deposition and might be insufficient to reflect the dependences of plants on animals and vice versa [@Alarcon2010; @King2013]. 
We therefore investigated the effect of calculating the dependences using visitation or pollination effectiveness and importance---two metrics more proximate to plant reproductive success (Figure \@ref(fig:fig-visitation-vs-deposition-vs-importance)).
We did this by comparing *(i)* the manageability of the community and *(ii)* the percentage of interactions that maintained the direction of dependence. 
To do that, we used data collected by @Ballantyne2015 from a low diversity pollination community at a dry lowland heathland in Dorset, UK  (50° 43.7'N 2° 07.2'W). 
First, deposition networks were quantified using the mean single visit deposition---the number of conspecific pollen grains effectively deposited on a virgin stigma during a single visit by a particular animal [@NeEman2010; @King2013; @Ballantyne2015]. 
Second, visitation networks were constructed counting the visits to flowers during Single Visit Depositions. 
Finally, pollinator importance networks were constructed as the product of pollinator efficiency and visit frequency.

@@

The choice of weighting used can also have an impact on at the species level.
Therefore we calculated *f~D~* and *f~S~* (the frequency at which each species is classified as a driver or superior node, respectively, within the set of all possible matchings in a network) and calculated it's correlation among all three weighting schemes.
Although visitation and deposition produce moderately different results, we found a very strong agreement between the order produced by visitation and importance (Table S2). 
Finally, we also investigated whether the asymmetry of mutual dependence, which defines the direction of control, was consistent among the three possible weighting schemes. We found again that the direction of the dominant dependence was maintained was consistent for 95\% of the interactions weighted by visitation or importance, the two most appropriate metrics for pollinator and plant dependence (Table S2).

@@

mapply(my_paste, correlations$r, correlations$P) %>%
  matrix(ncol = length(dir_net_b)) %>% magrittr::set_colnames(nam) %>%
  magrittr::set_rownames(nam) %>% `[<-`(lower.tri(.), lt) %>% 
  xtable::xtable(caption = "\\textbf{Agreement among network weighting schemes}. Spearman correlation coeffcients (with p-value) of species' $f_D$ and the percentage of interactions that share the direction of dependence obtained using the three weighting schemes and an unweighted scheme.") %>%
	print(comment = FALSE, include.rownames = TRUE, caption.placement = "top", booktabs = T)

@@

Altogether, evidence supports the idea that visitation is a suitable metric to estimate the mutual dependence of species pairs. First, it is directly related to pollinator foraging. Second, it produces results consistent, at least within our controllability framework, with plant reproductive success (as estimated by the importance metric).
